### 2024-07-25 paper.tex
say what a, b look like in case of general X

\subsection{A necessary condition for \(A_{X, a, b}\) to be nilpotent}
Suppose \(A := A_{X, a, b}\) is nilpotent.
Define the \emph{height} of a vector \(v \in \mathbb{C}^{m + n}\) as the smallest \(k \geq 0\) such that \(A_{X, a, b}^k v = 0\).
Clearly, for \(i \leq j\), we have that the height of \(f_i\) is geq the height of \(f_j\).
Thus, for each \(k \geq 0\), we must have \(A_{X, a, b}^k f_n \in \langle e_{ij} \rangle_{ij}\).
We have \(Af_n = \sum_{ij} b_{ij} e_{ij}\).
Then \(A^2 f_n = \sum_{ij} b_{ij} (e_{i, j - 1} + a_{ij} f_1)\), so we must have \(\sum_{ij} b_{ij} a_{ij} = 0\).
Similarly, since \(A^3 f_n \in \langle e_{ij}\rangle\), we see that \(\sum_{ij} b_{ij} a_{i, j - 1} = 0\).
Continuing in this way, we obtain that for all \(k \geq 0\),
\[\sum_{ij} b_{ij} a_{i,j - k} = 0.\]
In fact this is in fact a sufficient condition for \(A_{X, a, b}\) to be nilpotent, as can be seen from the characteristic polynomial of \(A_{X, a, b}\).
However, we will not do this calculation here, instead just showing that it is sufficient by giving a Jordan basis.

@@

\subsection{When \(X\) is a Jordan block}
Suppose \(X\) is a single Jordan block of size \(m\).
In this case we just write the basis of \(\mathbb{C}^m\) as \(e_1, ..., e_m\), and the above condition on \(a\) and \(b\) simply becomes
\begin{equation}
    \forall k \geq 0. \; \sum_{i = k + 1}^m b_i a_{i - k} = 0.
\end{equation}
This condition can be simplified even more.
\begin{lemma}
    The condition (1) holds iff there exist nonnegative \(m_1, m_2, m_3\), \(a_1, ..., a_{m_3}\), and \(b_1, ..., b_{m_3}\) satisfying the following conditions.
    \begin{itemize}
        \item \(m = m_1 + m_2 + m_3\)
        \item \(a = (a_1, ..., a_{m_1}, 0, 0, ..., 0)\)
        \item \(b = (0, 0, ..., 0, b_1, ..., b_{m_3})\)
        \item If \(m_1 \neq 0\), then \(a_{m_1} \neq 0\)
        \item If \(m_3 \neq 0\), then \(b_1 \neq 0\)
    \end{itemize}
\end{lemma}
\begin{proof}
    If \(a\) and \(b\) are of this form, then for every \(k \geq 0\) we have \(b_i a_{i - k} = 0\), so clearly the condition holds.
    \par Now suppose (1) holds.
    If \(a\) or \(b\) is zero, this is trivial.
    Otherwise, let \(m_1 = \max\{i : a_i \neq 0\}\), and let \(m_3 = \min\{i : b_i \neq 0\}\).
    We just need to show that \(m_1 < m_3\).
    For contradiction, suppose \(m_1 - m_3\) is nonnegative.
    Then by (1), \(0 = \sum_i b_i a_{i - (m_1 - m_3)} = b_1 a_{m_1}\), contradicting that \(b_1\) and \(a_{m_1}\) are nonzero.
\end{proof}

\subsection{In the case of general \(X\)}
In the case where \(X\) was a single Jordan block, it was helpful to have the condition that all the nonzero \(a\)'s were to the right of all the nonzero \(b\)'s.
In the general case, it is not clear that such a thing should be true.
And indeed it is not.
However, we will see that we can change basis of \(\mathbb{C}^m\) so that a similar thing is true.
\par Let \(c_{i',j'}^i \in \mathbb{C}\) be some arbitrary coefficients.
Given our Jordan basis \(e_{ij}\) of \(\mathbb{C}^m\), we can define a new Jordan basis \(e_{ij}'\) of \(\mathbb{C}^m\) by 
\[e_{i, \lambda_i}' = e_{i, \lambda_i} + \sum_{\{(i',j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c^i_{i',j'} e_{i',j'}.\]
Then we define
\[e_{i,j}' = X^{\lambda_i - j} e_{i, \lambda_i}' = e_{ij} + \sum_{\{(i', j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c^i_{i',j'} e_{i', j' - (\lambda_i - j)}.\]
Let us calculate the new value \(a'\) of \(a\) in this new basis.
That is, we want the values \(a_{ij}'\) that satisfy \(Ae_{ij}' = e_{i, j - 1}' + a_{ij}' f_1\).
Looking at the definition of \(e_{ij}'\) above, we see that
\[a_{ij}' = a_{ij} + \sum_{\{(i',j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c_{i',j'}^i a_{i', j' - (\lambda_i - j)}.\]
\par Now, for each \(i\), choose the coefficients \(c_{i',j'}^i\) to maximize the value \(\max\{j : (a_{i,1}', ..., a_{i, j}') = 0\}\).
In other words, we choose \(c_{i'j'}^i\) to zero out the largest possible prefix of \((a_{i, 1}, ..., a_{i, \lambda_i})\).

>>>>

\subsection{A necessary condition for \(A_{X, a, b}\) to be nilpotent}
Suppose \(A := A_{X, a, b}\) is nilpotent.
Define the \emph{height} of a vector \(v \in \mathbb{C}^{m + n}\) as the smallest \(k \geq 0\) such that \(A_{X, a, b}^k v = 0\).
Clearly, for \(i \leq j\), we have that the height of \(f_i\) is geq the height of \(f_j\).
Thus, for each \(k \geq 0\), we must have \(A_{X, a, b}^k f_n \in \langle e_{ij} \rangle_{ij}\).
We have \(Af_n = \sum_{ij} b_{ij} e_{ij}\).
Then \(A^2 f_n = \sum_{ij} b_{ij} (e_{i, j - 1} + a_{ij} f_1)\), so we must have \(\sum_{ij} b_{ij} a_{ij} = 0\).
Similarly, since \(A^3 f_n \in \langle e_{ij}\rangle\), we see that \(\sum_{ij} b_{ij} a_{i, j - 1} = 0\).
Continuing in this way, we obtain that for all \(k \geq 0\),
\begin{equation}\label{ab_nilp}\sum_{ij} b_{ij} a_{i,j - k} = 0.\end{equation}
In fact this is in fact a sufficient condition for \(A_{X, a, b}\) to be nilpotent, as can be seen from the characteristic polynomial of \(A_{X, a, b}\).
However, we will not do this calculation here, instead just showing that it is sufficient by giving a Jordan basis.

@@

\subsection{When \(X\) is a Jordan block}
Suppose \(X\) is a single Jordan block of size \(m\).
In this case we just write the basis of \(\mathbb{C}^m\) as \(e_1, ..., e_m\), and the above condition on \(a\) and \(b\) simply becomes
\begin{equation}
    \forall k \geq 0. \; \sum_{i = k + 1}^m b_i a_{i - k} = 0.
\end{equation}
This condition can be simplified even more.
\begin{lemma}
    The condition (1) holds iff there exist nonnegative \(m_1, m_2, m_3\), \(a_1, ..., a_{m_3}\), and \(b_1, ..., b_{m_3}\) satisfying the following conditions.
    \begin{itemize}
        \item \(m = m_1 + m_2 + m_3\)
        \item \(a = (0, 0, ..., 0, a_1, ..., a_{m_3})\)
        \item \(b = (b_1, ..., b_{m_1}, 0, 0, ..., 0)\)
        \item If \(m_1 \neq 0\), then \(a_{m_1} \neq 0\)
        \item If \(m_3 \neq 0\), then \(b_1 \neq 0\)
    \end{itemize}
\end{lemma}
\begin{proof}
    If \(a\) and \(b\) are of this form, then for every \(k \geq 0\) we have \(b_i a_{i - k} = 0\), so clearly the condition holds.
    \par Now suppose (1) holds.
    If \(a\) or \(b\) is zero, this is trivial.
    Otherwise, let \(m_1 = \max\{i : a_i \neq 0\}\), and let \(m_3 = \min\{i : b_i \neq 0\}\).
    We just need to show that \(m_1 < m_3\).
    For contradiction, suppose \(m_1 - m_3\) is nonnegative.
    Then by (1), \(0 = \sum_i b_i a_{i - (m_1 - m_3)} = b_1 a_{m_1}\), contradicting that \(b_1\) and \(a_{m_1}\) are nonzero.
\end{proof}

\subsection{In the case of general \(X\)}
In the case where \(X\) was a single Jordan block, it was helpful to have the condition that all the nonzero \(a\)'s were to the right of all the nonzero \(b\)'s.
In the general case, it is not clear that such a thing should be true.
And indeed it is not.
However, we will see that we can change basis of \(\mathbb{C}^m\) so that a similar thing is true.
\par Let \(c_{i',j'}^i \in \mathbb{C}\) be some arbitrary coefficients.
Given our Jordan basis \(e_{ij}\) of \(\mathbb{C}^m\), we can define a new Jordan basis \(e_{ij}'\) of \(\mathbb{C}^m\) by 
\[e_{i, \lambda_i}' = e_{i, \lambda_i} + \sum_{\{(i',j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c^i_{i',j'} e_{i',j'}.\]
Then we define
\[e_{i,j}' = X^{\lambda_i - j} e_{i, \lambda_i}' = e_{ij} + \sum_{\{(i', j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c^i_{i',j'} e_{i', j' - (\lambda_i - j)}.\]
Let us calculate the new value \(a'\) of \(a\) in this new basis.
That is, we want the values \(a_{ij}'\) that satisfy \(Ae_{ij}' = e_{i, j - 1}' + a_{ij}' f_1\).
Looking at the definition of \(e_{ij}'\) above, we see that
\[a_{ij}' = a_{ij} + \sum_{\{(i',j') : i' \leq i, j' \leq \lambda_i\} \setminus \{(i, j)\}} c_{i',j'}^i a_{i', j' - (\lambda_i - j)}.\]
\par Now, for each \(i\), choose the coefficients \(c_{i',j'}^i\) to maximize the value \(\max\{j : (a_{i,1}', ..., a_{i, j}') = 0\}\).
In other words, we choose \(c_{i'j'}^i\) to zero out the largest possible prefix of \((a_{i, 1}, ..., a_{i, \lambda_i})\).
\begin{lemma}
    The condition of \cref{ab_nilp} on \(a\) and \(b\) holds iff there exist nonnegative \(p_i, q_i, r_i\), \(\alpha_{i, j}\), \(\beta_{i, j}\) such that for each \(i\), the following conditions are satisfied.
    \begin{itemize}
        \item \(\lambda_i = p_i + q_i + r_i\)
        \item \((a_{i,1}, ..., a_{i,\lambda_i}) = (0, 0, ..., 0, \alpha_{i,1}, ..., \alpha_{i, r_i})\)
        \item \((b_{i, 1}, ..., b_{i, \lambda_i}) = (\beta_{i, 1}, ..., \beta_{i, p_i}, 0, 0, ..., 0)\)
        \item If \(r_i \neq 0\), then \(a_1 \neq 0\)
        \item If \(p_i \neq 0\), then \(b_{p_i} \neq 0\)
    \end{itemize}
\end{lemma}
\begin{proof}
